NumPy---Array Computation for Python

### 2020-01-05 body.tex
More work on outline

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
Stéfan van der Walt is the PI and manages four programmers working on the project.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767

% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received another grant for \$195,000 from the Chan Zuckerberg
Initiative at the end of 2019 with Ralf Gommers as the PI.
This grant, building on the technical work supported by the previous
grant, will focus on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
It will also provide support to OpenBLAS, on which NumPy depends for
accelerated linear algebra.

Finally, the project receives around 10K USD annually from Tidelift, which is
typically used to fund documentation writers and website developers.

@@

Given the complexity of the codebase and the massive number of projects depending
on it, large changes require careful planning and substantial work.
As the NumPy release manager in 2007, Jarrod Millman, adopted the notion
of enhancement proposals---used infrequently and informally until 2017 when,
to facilitate work expected from funded development, he formalized the
NumPy Enhancement Proposal (NEP) process.  NEPs are modeled after
Python Enhancement Proposals (PEPs) for ``proposing major new
features, for collecting community input on an issue, and for
documenting the design decisions that have gone into
Python''\footnote{\url{https://numpy.org/neps/nep-0000.html}}.
%Since then there have been 19 proposed NEPS---6 have been implemented,
%4 have been accepted and are being implemented, 4 are under
%consideration, 3 have been deferred or superseded, and 2 have been rejected
%or withdrawn.

The first activities organized for the new grant were a NEP development
sprint, a weekly community call, and a meeting to write a draft roadmap.
Since the roadmap was meant to represent community consensus on future
development, it was presented for comment at the annual SciPy
conference during a so-called ``Birds of a Feather'' session, attended
by more than 100 people.  One additional meeting was held to discuss
specific enhancements, such as the development of a new data-type
system.  The final roadmap was vetted by the community via online
discussion before being accepted.

@@

%Moreover, with the now prevalent need for distributed computing and the
%emergence of specialized hardware such as GPUs, many new usecases have
%arisen, which NumPy was not originally designed for.
With the recent infusion of funding and the newly formalized NEP process in
place we have been able to tackle a number of important large scale changes.
We highlight two of those below.

>>>>

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received a third grant for 200K USD from the Chan Zuckerberg
Initiative at the end of 2019.
This grant focuses on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
Finally, the project receives around 10K USD annually from Tidelift, which is
typically used to fund documentation writers and website developers.

The first activities organized for the new grants were a development planning
sprint, a weekly community call, and a meeting to write a draft roadmap.
Since the roadmap was meant to represent community consensus on future
development, it was presented for comment at the annual SciPy
conference during a so-called ``Birds of a Feather'' session, attended
by more than 100 people.  One additional meeting was held to discuss
specific enhancements, such as the development of a new data-type
system.  The final roadmap was vetted by the community via online
discussion before being accepted.

@@

@@

%Moreover, with the now prevalent need for distributed computing and the
%emergence of specialized hardware such as GPUs, many new usecases have
%arisen, which NumPy was not originally designed for.

With the recent infusion of funding and a clear process for coordinating with
the developer community, we have been able to tackle a number of important
large scale changes.
We highlight two of those below.
